### 2024-07-06 doc/Assignment2.qmd
Deleted unneccesary code and data

---
title: "Assignment II Analysis"
author: "Melisa Mazaeva"
date: "2024-07-12"
format: html
---

# Introduction
This document presents the analysis performed in Assignment 1, adapted to the open science workflow. Note that intermediate check steps have been omitted.

# Methodology
## Data Loading
```{python}
import pandas as pd
import numpy as np

# Load the data
static_data = pd.read_csv('data/external/wscp_static.txt', delimiter='\t')
panel_data = pd.read_excel('data/external/wscp_panel.xlsx')

# Display first few rows of the datasets
print(static_data.head())
print(panel_data.head())