### 2024-07-11 doc/presentation_mine.qmd
Created documentation presentation

---
title: "Assignment II: Adopting an Open Science Workflow"
author: "Melisa Mazaeva"
institute: "Accounting Reading Group – Summer 2024"
date: today
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: true
    toc: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\fontsize{7}{7}\selectfont}  # This sets the font size for verbatim environments
#- \input{beamer_theme_trr266.sty}
bibliography: references.bib
biblio-style: apsr
---

# Motivation

## Understanding Residual Income and Importance of P/B Ratios

- Residual income is a key metric in evaluating the performance of public firms.
- It helps in assessing the value creation for shareholders beyond the cost of capital.
- Price-to-Book (P/B) ratios are widely used in financial analysis to compare a firm’s market value to its book value.
- Grouping firms by P/B ratios allows for a detailed analysis of financial performance across different valuation levels.

## Project Objective

- The primary objective is to demonstrate a reproducible and collaborative research workflow by calculating and analyzing residual incomes and Price-to-Book (P/B) ratios. 
- The study aims to explore the residual income of U.S. public firms grouped by P/B ratios using external Worldscope datasets for the time range 1996 to 2015.
- The project replicates the study of @Penman_2013 that links current P/B ratios to future residual income.
- The analysis provides insights into the median residual income over time across different P/B groups, similar to the insights of Penman (2013).

## Project Relevance

- Understanding the distribution and trends of residual income can inform investment decisions and corporate strategies.
- It supports the broader objective of promoting transparency and reproducibility in empirical accounting research.

# Assumptions and Notes
The aim of Assignment I is to replicate a specific empirical table that involves calculating the residual income of US firms over a defined period and examining the relationship between residual income and the Price-to-Book (P/B) ratio. The replication process involves data preparation, cleaning, and normalization, followed by the application of statistical methods to compute and interpret financial metrics.

For Assignment I, I used the Python programming language to carry out the empirical analysis. Visual Studio Code was used as the Integrated Development Environment (IDE) for writing, debugging, and optimizing the Python code.

## Assumptions
- **Constant Cost of Equity:** Assumed at 8% ($r_e = 8\%$), a reasonable average estimate for the period 1996-2015 in the US (@Goedhart_Koller_Williams_2002; @Damodaran_2024).
- **Data Source:** Worldscope panel data includes Book Value of Equity (BVE), Market Value of Equity (MVE), and net income at fiscal year-end .
- **Year 0 Data:** Year-end data from the previous year used as the starting values for the following year.
- **Deflation Basis:** Residual income deflated by BVE at the end of the year before Year 0 for comparability.
- **P/B Values Range:** Restricted to 0-7 to exclude outliers, focusing on firms with stable valuations.

# Data - Replication Steps

Step 1: Data Loading and Merging

Step 2: Filter Data for US Firms and Relevant Years

Step 3: Identify Year 0 to Year 6 for Each Firm

Step 4: Calculate P/B Ratio for Each Firm in Year 0

+ Use the following formula to calculate the P/B ratio:
$$
P/B_t = \frac{\textit{Market value of equity}_t}{\textit{Book value of equity}_t}
$$

Step 5: Form P/B Groups

---

Step 6: Calculate Residual Income

+ The Worldscope panel data provides BVE, MVE, and net income values at the end of each fiscal year, confirmed by the *2020 Annual Report* (2021) of ISIN AN8068571086 [@Schlumberger_2021, p. 20].
+ Calculate the non-deflated residual income for each firm/year using the following formula:
$$
\text{Residual income}_t = \text{Net income}_t - r_e \times \text{Book value of equity}_{t-1}
$$

+ Calculate the deflated residual income for each firm/year using the following general formula:
$$
\text{Deflated residual income}_t = \frac{\text{Residual income}_t}{\text{Book value of equity}_{\text{Year 0 start}}}
$$

---

Step 7: Output the Replicated Table

```{python}
#| label: step1-7
#| echo: false
#| output: true

import pandas as pd
import numpy as np
import os
import yaml

# Function to get the path of the current file
def get_current_path():
    try:
        return os.path.dirname(os.path.abspath(__file__))
    except NameError:
        return os.getcwd()

# Define the relative path to the configuration file
current_path = get_current_path()
project_root = os.path.abspath(os.path.join(current_path, '..'))
config_path = os.path.join(project_root, 'config', 'config.yaml')

# Load the configuration from YAML file
with open(config_path, 'r') as file:
    config = yaml.safe_load(file)

# Extract paths from the config file
STATIC_DATA_PATH = os.path.join(project_root, 'data', 'external', 
                                'wscp_static.txt')
PANEL_DATA_PATH = os.path.join(project_root, 'data', 'external', 
                               'wscp_panel.xlsx')

# Load the data from Worldscope files into pandas DataFrames
static_data = pd.read_csv(STATIC_DATA_PATH, delimiter='\t')
panel_data = pd.read_excel(PANEL_DATA_PATH)

# Normalize column names to avoid issues with column naming conventions
static_data.columns = (
    static_data.columns.str.strip()
    .str.lower()
    .str.replace(' ', '_')
)

panel_data.columns = (
    panel_data.columns.str.strip()
    .str.lower()
    .str.replace(' ', '_')
)

# Merge data on 'isin' to include country information in panel_data
merged_data = panel_data.merge(
    static_data[['isin', 'country']], 
    on='isin', 
    how='left'
)

us_firms = merged_data[merged_data['country'] == 'UNITED STATES']  
filtered_data = us_firms[
    (us_firms['year_'] >= 1996) & 
    (us_firms['year_'] <= 2015)
].copy()

filtered_data['year_0'] = filtered_data.groupby('isin')['year_'].transform('min')  
filtered_data['relative_year'] = filtered_data['year_'] - filtered_data['year_0']  
filtered_data = filtered_data[filtered_data['relative_year'] <= 6]

year_0_data = filtered_data[filtered_data['relative_year'] == 0].copy()

# replace zero and very small BVE values with NaN
year_0_data['bve'] = year_0_data['bve'].replace([0, np.inf, -np.inf], np.nan)

# Use P/B Ratio formula to calculate it for each firm in Year 0
year_0_data['p/b'] = year_0_data['mve'] / year_0_data['bve']

# Remove rows with NaN P/B ratios
year_0_data = year_0_data.dropna(subset=['p/b'])

# Sort firms based on their P/B ratios for Year 0 in descending order
num_groups = 20  # define the number of groups
year_0_data = year_0_data.sort_values(
    'p/b', ascending=False
).reset_index(drop=True)

# Assign the groups so that the highest P/B ratios are in group 1
year_0_data['p/b_group'] = pd.qcut(
    year_0_data.index,
    num_groups,
    labels=range(1, num_groups + 1)
)

# Calculate median P/B value for each group
median_pb_values = year_0_data.groupby('p/b_group', observed=True)['p/b'].median()

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.merge(
    year_0_data[['isin', 'p/b_group']],
    on='isin',
    how='left'
)

# Fill missing P/B groups for years other than Year 0 with same group as Year 0
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()

cost_of_equity = 0.08  # assumption for cost of equity capital

# Create a copy of the merged data from Step 1 to keep all years of observations
us_full_data = merged_data[merged_data['country'] == 'UNITED STATES'].copy()

# Identify Year 0 for each firm, ensuring Year 0 is between 1996 and 2015
us_full_data['year_0'] = us_full_data.groupby('isin')['year_'].transform(
    lambda x: x[(x >= 1996) & (x <= 2015)].min()
)
us_full_data = us_full_data[us_full_data['year_0'].notna()] 
# Calculate the relative year and year_minus_1
us_full_data['relative_year'] = us_full_data['year_'] - us_full_data['year_0']
us_full_data['year_minus_1'] = us_full_data['year_0'] - 1

# Convert years to integer
us_full_data['year_0'] = us_full_data['year_0'].astype(int)
us_full_data['relative_year'] = us_full_data['relative_year'].astype(int)
us_full_data['year_minus_1'] = us_full_data['year_minus_1'].astype(int)

# Extract BVE for the year before Year 0 (relative_year = -1) for deflation
bve_deflation = (
    us_full_data[us_full_data['relative_year'] == -1][['isin', 'bve']]
    .rename(columns={'bve': 'bve_deflation'})
)

# Merge the bve_deflation values back to the filtered data
filtered_data = filtered_data.merge(bve_deflation, on='isin', how='left')

# Calculate the BVE of the previous year for each year
us_full_data['bve_prev'] = us_full_data.groupby('isin')['bve'].shift(1)
# Merge bve_prev back to filtered_data
filtered_data = filtered_data.merge(
    us_full_data[['isin', 'year_', 'bve_prev']],
    on=['isin', 'year_'],
    how='left'
)

# Function to calculate non-deflated residual income
def calculate_residual_income(row):
    return row['ninc'] - (cost_of_equity * row['bve_prev'])
# Apply the function to calculate non-deflated residual income for each row
filtered_data['residual_income'] = filtered_data.apply(
    calculate_residual_income, axis=1
)

# Function to deflate residual income
def deflate_residual_income(row):
    if pd.notna(row['bve_deflation']) and row['bve_deflation'] != 0:
        return row['residual_income'] / row['bve_deflation']
    else:
        return np.nan
# Apply the function to deflate residual income for each row
filtered_data['deflated_residual_income'] = filtered_data.apply(
    deflate_residual_income, axis=1
)

median_residual_income = filtered_data.groupby(
    ['p/b_group', 'relative_year'], observed=True
)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income = median_residual_income.apply(
    lambda col: col.map(
        lambda x: f"{x:.3f}" if pd.notna(x) else "NaN"
    )
)
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

# Add the median P/B values column to the output table
median_residual_income.insert(0, 'P/B', median_pb_values)

# Display the median residual income table
print(
    "Median Residual Income by P/B Group and Relative Year:\n", 
    median_residual_income
) 
```

---

Step 8: Modification - Exclude Extreme P/B Values

```{python}
#| label: step8
#| echo: false
#| output: true

# Filter out extreme P/B values that are negative and higher than 7
filtered_year_0_data = year_0_data[
    (year_0_data['p/b'] > 0) & (year_0_data['p/b'] <= 7)
]

# Recalculate the P/B groups based on the filtered data
filtered_year_0_data = filtered_year_0_data.sort_values(
    'p/b', ascending=False
).reset_index(drop=True)

filtered_year_0_data['p/b_group'] = pd.qcut(
    filtered_year_0_data.index, num_groups, labels=range(1, num_groups + 1)
)

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.drop(columns=['p/b_group'])
filtered_data = filtered_data.merge(
    filtered_year_0_data[['isin', 'p/b_group']], 
    on='isin', 
    how='left'
)
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()

# Calculate median P/B values again
median_pb_values = filtered_year_0_data.groupby(
    'p/b_group', 
    observed=True
)['p/b'].median()

# Calculate the median deflated residual income for each group and year
median_residual_income_filtered = filtered_data.groupby(
    ['p/b_group', 'relative_year'], 
    observed=True
)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income_filtered = median_residual_income_filtered.apply(
    lambda col: col.map(lambda x: f"{x:.3f}" if pd.notna(x) else "NaN")
)
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

# Add the median P/B values column to the output table
median_residual_income_filtered.insert(0, 'P/B', median_pb_values)

# Display the median residual income table
print(
    "Median Residual Income by P/B Group and Relative Year:\n",
    median_residual_income_filtered
)
```

# Conclusion

## Summary of Findings

- Replicated an empirical table from Penman (2013), showcasing similar trends.
- The project structure demonstrated the effectiveness of an open science and collaborative workflow.
- This repository can be cloned or forked for further projects on residual income analysis.

## Key Insights

- From analysis of residual incomes and Price-to-Book (P/B) ratios of U.S. public firms: 
    - High P/B groups initially exhibit higher residual earnings, which decline over time. 
    - Low P/B groups consistently show negative residual earnings.
- Residual income analysis by P/B ratios offers valuable insights that can promote better investment and corporate strategy decisions.

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent